Double White Dwarf Mergers with CASTRO\\ I. Methodology and Code Verification

### 2014-12-22 papers/wdmerger_I/paper.tex
Kepler's third law was inverted.
Added a description of the problem setup for the KH tests to the paper.

For a single star simulation, the star is simply placed at the center of the computational domain, which we take to be the origin. For a binary star simulation, we take as parameters the mass of the two white dwarfs and the initial orbital period $T$. Using Kepler's third law and assuming a circular orbit, we can then work out the orbital separation $a$:
\begin{equation}
  a = \left(\frac{4\pi^2 T^2}{GM}\right)^{1/3}.
\end{equation}
Here $M = M_P + M_S$ is the total mass of the system, where $M_P$ is the specified \textit{primary} mass and $M_S$ is the specified \textit{secondary} mass. The primary WD will always be on the left side of the computational domain for our simulations, and is more massive than the secondary. This reflects the usual terminology in the literature where the primary WD is the accretor and the secondary is the donor. The center of mass is located at the center of the computational domain, and the stars lie along the $x$ axis, so that the primary's center of mass is located at $x = -(M_S / M)\, a$ and the secondary's center of mass is located at $x = (M_P / M)\, a$.

@@

>>>>

For a single star simulation, the star is simply placed at the center of the computational domain, which we take to be the origin. For a binary star simulation, we take as parameters the mass of the two white dwarfs and the initial orbital period $T$. Using Kepler's third law and assuming a circular orbit, we can then work out the orbital separation $a$:
\begin{equation}
  a = \left(\frac{GM T^2}{4\pi^2}\right)^{1/3}.
\end{equation}
Here $M = M_P + M_S$ is the total mass of the system, where $M_P$ is the specified \textit{primary} mass and $M_S$ is the specified \textit{secondary} mass. The primary WD will always be on the left side of the computational domain for our simulations, and is more massive than the secondary. This reflects the usual terminology in the literature where the primary WD is the accretor and the secondary is the donor. The center of mass is located at the center of the computational domain, and the stars lie along the $x$ axis, so that the primary's center of mass is located at $x = -(M_S / M)\, a$ and the secondary's center of mass is located at $x = (M_P / M)\, a$.

@@

Following \cite{robertson:2010}, we set up a Kelvin-Helmholtz test in the following way. The problem domain runs from 0 to 1 in both the $x$ and $y$ directions. Though this is a two-dimensional test, we run CASTRO in 3D and replicate the problem along 8 cells in the $z$ direction; the physical size of these zones is determined so that $dz = dx = dy$, a requirement at present in CASTRO. The problem involves a fluid slab of density $\rho_1 = 2.0$ traveling rightward in the $x$-direction at velocity $v_1 = 0.5$, sandwiched by a fluid of density $\rho_2 = 1.0$ traveling leftward at velocity $v_2 = -0.5$. The density gradient is in the $y$ direction, so this creates a velocity shear along the interface between the fluids. The density and velocity distribution on the computational domain are given by:
\begin{align}
  \rho &= \rho_1 + R(y)\left[\rho_2 - \rho_1\right] \\
  v_x  &= v_1 + R(y)\left[v_2 - v_1\right] \\
  v_y  &= v_{\text{bulk}} + v^\prime
\end{align}
Here $R(y)$ is a ramp function that describes the transition between the two fluids, while $v_{\text{bulk}}$ is the bulk motion of the fluid in the $y$ direction and $v^\prime$ is the velocity perturbation that seeds the instability. The problem will be established for two sets of initial conditions (ICs), which we follow \citeauthor{robertson:2010} in calling ICs A and B. They differ in their ramp function ($R_A$ and $R_B$ respectively), as well as the initial perturbation ($v^\prime_A$ and $v^\prime_B$ respectively), and the frequency of the perturbation ($n_A = 4$ and $n_B = 2$):
\begin{align}
  R_A &= \begin{cases} 0 & |y - 0.5| > 0.25 \\ 1 & |y - 0.5| < 0.25 \end{cases} \\
  R_B &= \Big\{\left[1 + \text{exp}(-2(y-0.25)/\Delta_y)\right]\left[1 + \text{exp}(2(y-0.75)/\Delta_y)\right]\Big\}^{-1} \\
  v^\prime_A &= w_0\, \text{sin}\left(n_A\, \pi\, x\right) \left\{\text{exp}\left[-\frac{(y-0.25)^2}{2\sigma^2}\right] + \text{exp}\left[-\frac{(y-0.75)^2}{2\sigma^2}\right]\right\} \\
  v^\prime_B &= w_0\, \text{sin}\left(n_B\, \pi\, x\right).
\end{align}
Here $w_0 = 0.1$ is the scale of the velocity perturbation, $\sigma = 0.05/\sqrt{2}$ controls the width of the Gaussian for IC A, and $\Delta_y$ is the transition distance scale for the smooth ramp of IC B. The pressure everywhere is set to $p = 2.5$, and we run this with a gamma-law equation of state set to $\gamma = 5/3$. Plotfiles are generated every 0.01 seconds, and the problem is run until $t = 2$.

We run the problem for $v_\text{bulk} = [0, 1, 3, 10, 30, 100]$, and for each set of initial conditions run the problem at resolutions of $64^2$, $128^2$, $256^2$, $512^2$. In addition, for each initial condition we run a simulation at the very high resolution of $2560^2$ for the stationary problem only. This serves as a reference solution to gauge the extent to which the bulk flow affects the development of the fluid instability.
